Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-03 paper/paper.tex
Add plots for wavefields
added some compression blurp

introduction on compression techniques - TODO Jan
Compression is an obvious addition to the tools applied to the specific data access pattern

One possible way of using compression to speed up full-waveform inversion or any adjoint-based
optimization is by simply compressing the field data before storing it and decompressing it when it
is required again. If the memory available is sufficient to store the complete forward data after
compression, the use of compression effectively replaces the use of revolve-based checkpointing.
This scenario has been discussed in different studies before - some of which we cover in the next
section. This paper focusses on a broader region, of which this scenario is just one part. In particular,
we discuss the scenarios where the forward field can not fit in memory even after compression, where
compression and Revolve would be used in conjunction.

>>>>

Another strategy to reduce the memory footprint of scientific applications is
data compression. Well-known general purpose data compression packages such as
gzip TODO cite or 7zip TODO cite, and compression for video and image
data such as JPEG-2000 TODO cite have been used for scientific data. More
recently, special purpose compression algorithms for scientific data have been
developed, such as ZFP or SZ TODO cite.

Lossless algorithms guarantee that the exact original data can be
recovered during decompression, whereas lossy algorithms introduce an error, but
often guarantee that the error does not exceed certain absolute or relative
error metrics. Typically, lossy compression is more effective in reducing the
data size, or in other words, offer a greater compression ratio. Most popular
compression packages offer various settings that allow a tradeoff between
compression ratio, accuracy, and compression and decompression time. In this
work we use the ZFP package to perform lossy compresssion.

It is worth noting that another popular data reduction strategy is to typecast
values into a lower precision format, for example, from double precision to
single precision TODO cite. This can be seen as a computationally cheap lossy
compression algorithm with a compression ratio of $2$.

Perhaps counterintuitively, compression can not only reduce the memory
footprint, but also speed up an application. Previous work has observed that the
compression and decompression time can be less than the time saved from the
reduction in data that needs to be communicated across MPI nodes or between a
GPU and a host computer TODO cite.

Another possible way of using compression to speed up full-waveform inversion or any adjoint-based
optimization is by simply compressing the field data before storing it and decompressing it when it
is required again. If the memory available is sufficient to store the complete forward data after
compression, the use of compression effectively replaces the use of revolve-based checkpointing.
This scenario has been discussed in different studies before - some of which we cover in the next
section. This paper focusses on a broader region, of which this scenario is just one part. In particular,
we discuss the scenarios where the forward field can not fit in memory even after compression, where
compression and Revolve would be used in conjunction.
